Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-03-31 Manuscript/manuscript.Rmd

Another aspect of TAD boundary modelling not currently being explored are the various feature engineering methods used to build the feature space of models. Different techniques can be employed to measure the relationship between TAD boundaries and genomic annotations. Typical approaches for modeling the association between TAD boundaries and epigenomic elements include overlap counts or percents [@mourad2016computational; @hong2017computational; @huang2015predicting]. Counts refer to the total number of instances where a genomic annotation overlaps with a genomic region of interest, while percents refers to the percent of overlap that occurred (usually in terms of base pairs). The use of distances from TAD boundaries to regions of various genomic features, measured in the number of bases between them, has largely been ignored. However, these distance characteristics have the potential to offer a more accurate spatial representation of which genomic features are associated with TAD boundaries. Likewise, any direct comparisons of the different feature engineering techniques remain unresearched.

>>>>

Another underexplored aspect of modeling the association between TAD boundaries and genomic annotations is the various types of relationships between them. Typical approaches include overlap counts or percents [@mourad2016computational; @hong2017computational; @huang2015predicting]. Counts refer to the number of times a genomic annotation directly overlaps with a TAD boundary region, e.g., a TAD boundary may overlap with several ChIP-seq peaks for a given transcription factor. Percents refers to the percent of overlap of a genomic annotation with a TAD boundary region, e. g., a 60% of a TAD boundary may overlap genomic region annotated as heterochromatin.

While direct overlaps represent the most straightforward measure of functional associations, they do not account for indirect relationships shown to be one of the major driving force of genome regulation [??? REFS]. The most prominent example maybe the enhancer-promoter relationships, with enhancers are frequently located hundreths or thousands bases away from their target promoters [??? REFS]. The use of distances between TAD boundaries and genomic annotations has largely been ignored. However, these distance features have the potential to offer a more accurate spatial representation of which genomic features are associated with TAD boundaries.
